Supporting Information: Improving cropland maps through tight integration of human and machine intelligence

### 2021-02-24 inst/manuscript/manuscript.Rmd
First section of Discussion with outline

We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. These factors make Ghana emblematic of the substantial challenges facing cropland classification over the broader region.

@@

Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @ShiCirrusCloudsThatAffect2020].

@@

In the second step, a Sobel filter is then applied to the green, red, near-infrared mean-shifted bands and probability map , and an combined edge image is computed using the sum of these four edge images for the dry season only. A compact watershed algorithm [@NeubertCompactWatershedPreemptiveSlic2014] is then run on the weighted edge image, with a high level of segmentation specified. In this case, we specified **6400** segments per tile.

@@

In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@VisvalingamLineGeneralisation1993]. Neighboring polygon that touch along tile boundaries are merged.

To assess the accuracy of the final segmented boundaries, we used a two-step approach. First, we assessed the overall thematic accuracy of the resulting classification against our map reference data. Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map. [**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@YeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery. On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).

@@

Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table 2). This overall level of accuracy varied within four smaller zones (Figure S9A), ranging from 83.3% in Zone 1 (comprised of AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16). Examining the accuracies of the individual classes, the Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1. The User's accuracy for cropland was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1. Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%, with the lowest accuracies for this class found in Zone 1 (Producer's = 89.3% and User's = 87.7%, respectively).

The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table 2). In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map, indicating that the segmentation step helped reduce omission error, but substantially increased false positive errors.

@@

\begin{table}
\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions; top 5 rows) and the segmented map (bottom 5 rows). Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country. The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error. }
\includegraphics[width = 18cm]{figures/table2.png}
\label{tab:aezaccuracy}
\end{table}

@@

# Discussion

We present here a method for developing high resolution maps that detail the characteristics of smallholder-dominated croplands, providing information on their geometries and distribution of <5 m resolution, and representing a single agricultural year.

The results 
Bias in mean sizes likely reflect agglomerations of contiguous smallholder fields, and the difficulty of segmenting indistinct adjacent boundaries. Such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).

@@

- Left unquantified is how much training on base maps matters
- Image quality:
    - why mean-weighting not median composite (artifacts) (maybe methods)
    - impact on accuracy
- Assessment of segmentation quality
- Next steps: 
    - Annually updated cropland maps
    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours. 
    - Feng et al (2020) importance of temporal information (backs up Debats et al).
      - temporal information makes labelling harder though

Improvements:    
- Replace Random Forest with deep learning model. Can still retain the active architecture, but instead use new sites to improve a pre-trained model
- Map along agroecozone boundaries, rather than random blocks

>>>>

We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. These factors make Ghana emblematic of the substantial challenges facing cropland mapping over the broader region.

@@

Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]

@@

In the second step, a Sobel filter is then applied to the green, red, near-infrared mean-shifted bands and probability map , and an combined edge image is computed using the sum of these four edge images for the dry season only. A compact watershed algorithm [@neubertCompactWatershedPreemptive2014] is then run on the weighted edge image, with a high level of segmentation specified. In this case, we specified **6400** segments per tile.

@@

In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@visvalingamLineGeneralisationRepeated1993]. Neighboring polygon that touch along tile boundaries are merged.

To assess the accuracy of the final segmented boundaries, we used a two-step approach. First, we assessed the overall thematic accuracy of the resulting classification against our map reference data. Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map. [**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@yeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery. On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).

@@

Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table \href{tab:mapaccuracy}). This overall level of accuracy varied within four smaller zones (Figure S9A), ranging from 83.3% in Zone 1 (comprised of AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16). Examining the accuracies of the individual classes, the Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1. The User's accuracy for cropland was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1. Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%, with the lowest accuracies for this class found in Zone 1 (Producer's = 89.3% and User's = 87.7%, respectively).

The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table \href{tab:mapaccuracy}). In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map, indicating that the segmentation step helped reduce omission error, but substantially increased false positive errors.

@@

\begin{table}
\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions; top 5 rows) and the segmented map (bottom 5 rows). Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country. The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error. }
\includegraphics[width = 18cm]{figures/table2.png}
\label{tab:mapaccuracy}
\end{table}

@@

# Discussion
```{r, echo=FALSE}
# from lesiv et al (2019) SI field sizes for Ghana
lesiv_fsizes <- tibble(n = c(6, 7, 25, 71, 154), 
                       sz = c(100, mean(c(100, 16)), mean(c(2.56, 16)), 
                              mean(c(0.64, 2.56)), 0.32))
lesiv_est <- sum(with(lesiv_fsizes, (n / sum(n))  * sz))

aoi_nofor <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
  summarize(weighted.mean(fsize, ntiles)) %>% pull() %>% round(., 2)
aoi_nofor_adj <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
  summarize(weighted.mean(fsizeadj, ntiles)) %>% pull() %>% round(., 2)
```

These results demonstrate our system's ability to create high resolution maps of the distribution and characteristics of cropland across a diverse and hard-to-map agricultural systems, at country-scales and at sub-annual to annual time steps. Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our maps of cropland in 2018, which cover 16.1-23.2% of the mapped area (Table \href{tab:mapaccuracy}), and yield a sample-corrected estimate `r round(c(42359) / (42359 + 204940) * 100, 1)` to `r round(c(43233) / (43233 + 202856) * 100, 1)`%, which is lower than both prior estimates.

In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \href{fig:mainmap}, S12). Prior efforts to map such agricultural characteristics have often focused on estimating farm rather than field size using census datasets [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004]. Efforts to map fields in small-scale agricultural systems have either been field-based [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small areas [e.g. @forkuorIntegrationOpticalSynthetic2014]. The most notable efforts to estimate field sizes were two global initiatives that used crowdsourced volunteers to categorize five broad field size classes based on their interpretations of imagery sampled from virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015]. The most recent of these [@lesivEstimatingGlobalDistribution2019] provided country-specific results for Ghana (n = 263) that return an average field size of `r round(lesiv_est, 2)` ha (calculated as the weighted mean of the class counts and the middle values of the four smallest size classes and the lower bound of the largest size class). This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \href{tab:sizentab}), but is closer to the mean segment size (`r aoi_nofor` ha) across AOIs 1-9, 12, and 15, which is where most of their sample appears to have been collected. However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha are much smaller.

## Performance of the system
Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors. The overall map accuracies are good (86.7-88%, Table \href{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence). However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map, indicating that both commission and omission errors were above 30% for this class. The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%). These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017]

Although the raw maps are biased, data from the labelling platform provides information that can be used to correct for these discrepancies.
The detail, large extent, and temporal precision of the map were enabled by the system's ability to process small-sat data, which currently provide the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with the frequency necessary to create clean imagery in areas of frequent cloud cover across multiple crop growing periods, which is important for distinguishing cropland from non-cropland [@Debatsgeneralizedcomputervision2016]. This map represents an capability represents an advance over Although free and providing higher radiometric quality and depth, a moderate-resolution sensor such as Sentinel-2 may be too infrequent to successfully penetrate clouds in an area such as southern Ghana (meaning that dry season imagery is likely to be all that is available), while lacking the spatial resolution necessary for labelling and classifying smaller fields.

A possible reason for the discrepancy between earlier maps and ours is that those earlier maps include more areas of active or abandoned cropland in their estimates.

using a mapping system that is designed to work with high spatio-temporal resolution small-sat data.

- Map at large scale using PlanetScope, one of the first (any citations). 
  - Done using small sat data for a single agricultural year, for two primary seasons
  - Follows recommended best practice for training ML algorithms

- Takes step towards answering important questions about the characteristics of fields and how many there are. 
  - Provide insights into the characteristics of croplands
  - Discussion about the characteristics of croplands: estimates of field sizes
  - What is says about Ghana's croplands.

## Performance of the system
- Pixel-wise system, reasonable overall accuracy, but low performance on the agricultural class
- Low performance primarily due to several factors:
  - Model simplicity and scale of compute
  - Challenge of labelling task - uncertainty in labels
    - Should also be noted substantial uncertainty in identifying fields (map reference uncertainty)
  - Difficulty of getting clean imagery in highly cloudy areas
- Design of system helped to mitigate these factors:
  - Consensus labels minimize error substantially
  - Active learning selects cites more efficiently, some evidence it improves more efficiently and overall relative to random selection. Less marked gain here, but doing this at much larger scale than Debats
  - Multi-season imagery helps improve it    
    - Feng et al (2020) importance of temporal information (backs up Debats et al).
    - temporal information makes labelling harder though
  - Image quality:
    - Developed our own mean-weighting not median composite (artifacts) (maybe methods)
    - impact on accuracy--residual cloud issues
    - Important to develop the method because best-on-top method of Planet's basemaps leaves lots of residual cloud in key seasons, and require large volumes of imagery in highly cloudy areas such as southerwestern Ghana.

- Segmentation accuracy issues:
  - Over-merging
  - Bias in mean sizes likely reflect agglomerations of contiguous smallholder fields, and the difficulty of segmenting indistinct adjacent boundaries. Such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).
  - Importance of having high quality probabilities underlying for filtering segments

## Looking forward 
- Sets up capability for repeated annual mapping of croplands to understand critical issues of how these systems are changing (sizes and shapes of croplands, where they are expanding) (Jayne papers, recent agricultural expansion paper)
- Key components will be useful in other applications and regions
- Provides means for improving crop type and crop yield estimates
- Improvements to help realize these capabilities: promise of CNNs (cite Waldner paper, others), for segmentation to realize this. This systems is well designed to adapt to it, and capabilities for assessing various aspects of label quality, e.g. edge accuracy and fragmentation accuracy, can help improve quality of labels designed to emphasize edges
  - Can still retain the active architecture, but instead use new sites to improve a pre-trained model
- Map along agroecozone boundaries, rather than random blocks  
- Another interesting question to answer: whether guided approach to site selection would result in more informative sample for CNNs also.

@@

- Left unquantified is how much training on base maps matters

- Assessment of segmentation quality
- Next steps: 
    - Annually updated cropland maps
    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.

### 2021-02-24 inst/manuscript/si.Rmd
First section of Discussion with outline

@@

The average Bayes Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}). The three initial clusters include the second mapping of Cluster 2.

@@

```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayes Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines. The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_label_risk_hists.png')
```

>>>>

#### Worker's assignment evaluation
We assigned a score for each workers' assign; the score history determines the weights of each workers for creating a consensus label for each 0.005 degree resolution grid. The score for a particular assignment *i* is defined as a weighted sum of five accuracy metrics:

$$ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}$$

Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.

*I* is "inside the box" accuracy which is defined as a proportion of the area correctly mapped within a 0.005 degree resolution grid (*I$_c$*) over total "inside the box" area for this grid (*I$_t$*).

$$ I=\frac{I_c}{I_t}$$

*O* is "outside the box" accuracy which refers to a proportion of the field area correctly mapped outside the grid over total "outside the box" region (*O$_t$*, the region within the bounding box of the workers' polygons but not within 0.005 degree resolution grid).

$$ O=\frac{O_c}{O_t}$$

*F* is the fragmentation accuracy which is defined as a proportion of matched polygon number (*N$_m$*, the number of the workers' polygons that has at least 50% of its region overlapped with a reference polygon) over total workers' polygon number (*N$_t$*).

$$ F=\frac{N_m}{N_t}$$

*E* is the average edge accuracy for all pairs of matched workers and refernece polygons; the edge accuracy for a single pair is defined as the length of 'correctly mapped edges' (*L$_c$*, the partial boundary of a workers' polygon that are within a three-planet-pixel buffer region of the matched reference polygon boundary) over the total boundary length of its matched reference polygon (*L$_t$*).

$$ E=\frac{L_c}{L_t}$$

*C* is the categorical accuracy, i.e., a proportion of the area that has been correctly labeled with field category within intersected regions between worker's and reference polygons (*T$_c$*) over the total intersecting area (*T$_t$*).

$$ C=\frac{T_c}{T_t}$$

@@

The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}). The three initial clusters include the second mapping of Cluster 2.

@@

```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines. The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_label_risk_hists.png')
```
